### 2015-09-28 paper.tex

\documentclass[conference]{IEEEtran}
\usepackage{pifont}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{mybeamer}
%\usepackage{draftwatermark}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

\newcommand{\eg}{\mbox{{\em e.g.}}}
\newcommand{\ie}{\mbox{{\em i.e.}}}
\newcommand{\viz}{\mbox{{\em viz.}}}

\title{\Large {\bf Protocol-Guided Analysis of Post-silicon
  Traces Under Limited Observability}}

%% \author{\large Hao Zheng$^1$, Yuting Cao$^1$, Sandip Ray$^2$, Jin Yang$^2$ \\
%% $^1$Dept. of Computer Science and Eng., University of South Florida, Tampa, Fl 33620. USA. \\
%% $^2$Strategic CAD Labs, Intel Corporation, Hillsboro, OR 97124.  USA.}

\begin{document}

\maketitle

\begin{abstract}
We consider the problem of reconstructing system-level
behavior of an SoC design from a partially observed signal
trace.  Solving this problem is a critical activity in
post-silicon validation, and currently depends primarily on
human creativity and insight.  We provide algorithms to
automatically infer system-level transactions from
incomplete, ambiguous, and noisy trace data, together with a
measure of confidence.  We demonstrate the approach on
illustrative system-level protocols for SoC models developed
in SystemC as well as in an FPGA environment.
\end{abstract}

\section{Introduction}

Post-silicon validation makes use of pre-production silicon
integrated circuit (IC) to ensure that the fabricated system
works as desired under actual operating conditions with real
software.  Since the silicon executes at target clock speed,
post-silicon executions are billions of times faster than
RTL simulations, and even provide speed-up of several orders
of magnitude over other pre-silicon platforms (\eg, FPGA,
system-level emulation, etc.).  This makes it possible to
explore deep design states which cannot be exercised in
pre-silicon, and identify errors missed during pre-silicon
validation and debug.  Post-silicon validation is a critical
component of the design validation life-cycle for modern
microprocessors and SoC designs.  Unfortunately, it is also
a highly complex component, performed under aggressive
schedules and accounting for more than $50\%$ of the overall
design validation cost.  Consequently, it is crucial to
develop techniques for streamlining and automating
post-silicon validation activities.

A key component of post-silicon validation of SoC designs is
to correlate traces from silicon execution with the intended
system-level transactions.  An SoC design is typically
composed of a large number of pre-designed hardware or
software blocks (often referred to as ``intellectual
properties'' or ``IPs'') that coordinate through complex
protocols to implement the system-level behavior.  Any
execution trace of the system involves a large number of
interleaved instances of these protocols.  For example,
consider a smartphone executing a usage scenario where the
end-user browses the Web while listening to music and
sending and receiving occasional text messages.  Typical
post-silicon validation use-cases involve exercising such
scenarios.  An execution trace would involve activities from
the CPU, audio controller, display controller, wireless
radio antenna, etc., reflecting the interleaved execution of
several communication protocols.  On the other hand, due to
observability limitations, only a small number of
participating signals can be actually traced during silicon
execution.  Furthermore, due to electrical perturbations,
silicon data can be noisy, lossy, and ambiguous.
Consequently, it is non-trivial to identify all
participating protocols and pinpoint the interleaving that
results in an observed trace.

In this paper, we consider the problem of reconstructing
protocol-level behavior from silicon traces in SoC designs.
Given a collection of system-level communication protocols
and a trace of (partially observed) hardware signals, our
approach infers, with a certain measure of confidence, the
protocol instances (and their interleavings) being exercised
by the trace.  The approach is based on a formalization of
system-level transactions via labeled Petri-Nets, which are
capable of describing sequencing, concurrency, and choices
over system events.  We develop algorithms to infer
system-level transactions from traces with missing, noisy,
and ambiguous signal values.
% together with an estimate of confidence on the inference.  
We demonstrate our approach on
two SoC case studies: a pedagogical SystemC prototype and a
more realistic implementation constructed within the GEM5
environment~\cite{Binkert2011}.

\begin{figure*}[tb]
\begin{center}
\begin{tabular}{cc}
\begin{minipage}{3in}
\includegraphics[height=2in]{figures/bpmn-flow-ex}
\end{minipage}
&
\begin{minipage}{3in}
\resizebox{2.8in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(t1)	at	($(p1) + (0,-1.1)$) {$t_1:\langle {\tt Driver:Device:Load} \_fw\rangle$};
	\node[circ] 	(p2) 		at ($(t1) + (0,-1.1)$) {$p_2$};
	\node[rectbox]	(t2)	at	($(p2) + (0,-1.1)$) {$t_2:\langle {\tt Device:CE:Auth\_req} \rangle$};
	\node[circ] 	(p3) 		at ($(t2) + (0,-1.1)$) {$p_3$};
	\node[rectbox]	(t3)	at	($(p3) + (0,-1.1)$) {$t_3:\langle {\tt CE:Device:Auth\_resp}\rangle$};
	\node[circ] 	(p4) 		at ($(t3) + (0,-1.1)$) {$p_4$};
	\node[rectbox]	(t4)	at	($(p4) + (-3,-1.1)$) {$t_4:\langle {\tt Device:Driver:Report} \rangle$};
	\node[rectbox]	(t5)	at	($(p4) + (3,-1.1)$) {$t_5:\langle {\tt Device:CE:Ack} \rangle$};
	\node[circ,fill=black] 	(term1) 		at ($(t4) + (0,-1.1)$) {\textcolor{white}{$p_5$}};
	\node[circ,fill=black] 	(term2) 		at ($(t5) + (0,-1.1)$) {\textcolor{white}{$p_6$}};

	% Edges
	\draw[->,weight3]	(p1) -- (t1);
	\draw[->,weight3]	(t1) -- (p2);
	\draw[->,weight3]	(p2) -- (t2);
	\draw[->,weight3]	(t2) -- (p3);
	\draw[->,weight3]	(p3) -- (t3);
	\draw[->,weight3]	(t3) -- (p4);
	\draw[->,weight3]	(p4) -- ($(p4)+(-3,0)$) -- (t4);
	\draw[->,weight3]	(p4) -- ($(p4)+(3,0)$) -- (t5);
	\draw[->,weight3]	(t4) -- (term1);	
	\draw[->,weight3]	(t5) -- (term2);

\end{tikzpicture}
}
\end{minipage}
\\
(a) & (b)
\end{tabular}
\caption{(a) A graphical representation of a SoC firmware
  load protocol~\cite{Krstic14HOST}.  (b) LPN formalization.
  Each event has a form of $\langle {\tt src, dest, cmd}
  \rangle$ where ${\tt cmd}$ is a command sent from a source
  component ${\tt src}$ to a destination component ${\tt
    dest}$. The solid black places without outgoing edges
  are {\em terminals}, which indicate termination of
  protocols represented by the LPNs.}
\label{flow-spec-ex}
\end{center}
\end{figure*}

\section{Background}

\subsection{SoC Protocols and Post-silicon Trace Analysis}

An SoC design involves integration of a number of IPs that
communicate through complex protocols.  Such system-level
protocols are typically specified in architecture documents
as message flow diagrams.  As illustration,
Fig~\ref{flow-spec-ex}(a) shows one such diagram for a
protcol to authenticate and load a firmware during system
boot for firmware upgrade.  During validation, the system
under debug (SUD) exercises some complex system-level
use-case which involves interleaved execution of possibly a
large number of such protocols.  A trace of a small number
of hardware signals is then shipped off-chip analysis.  The
off-chip analysis includes two broad phases: (1)~trace
abstraction, and (2)~trace interpretation.  Trace
abstraction maps the hardware trace into higher-level
architectural constructs, \eg, messages, operations, etc.: a
message such as {\tt Authorization request} may be
implemented in hardware through a Boolean or temporal
combination of specific hardware signals in the NoC fabric
between {\tt Device} and {\tt CE}, \eg, as a sequence
containing a header, a specific value of a sequence of data
words, etc.  We will refer to such architectural constructs
as {\em protocol events} or {\em flow events}.  Note that
due to limited observability, it may not be possible to map
a given set of (observed) hardware signals uniquely to a
flow event.  Finally, the trace may a result from several
instances of the same protocol executing concurrently, \eg,
a firmware authentication protocol may be invoked when
another instance of the protocol has not completed.

Trace interpretation entails mapping flow events created
during trace abstraction to system-level protocols in order
to identify the set of protocol instances (and
interleavings) responsible for creating the observed
behavior.  The trace may identify a problem in the protocols
themselves, \eg~an interleaving of some protocol executions
may lead to an unexpected message being sent or cause the
system to crash.  More commonly, one finds a bug in the {\em
  implementation} of the protocol, \ie, a trace inconsistent
with any possible interleaving of the protocol executions.
Identifying these problems involves significant human
expertise, and can often take days to weeks of effort.

%% In a typical validation setting, the system under debug
%% (SUD) is executed in a test environment until it is
%% terminated by the test environment or the system crashes
%% due to a failure.  During the execution, a trace on a
%% small number of observable signals is streamed off the
%% chip for debugging.

\subsection{Labeled Petri-Nets}

Labeled Petri-nets (LPN) is a formalization of state
transition systems that is capable of describing sequencing,
concurrency, and choices.  Fig.~\ref{flow-spec-ex}(b)
illustrates how to use LPN to formalize protocols.
Formally, an LPN is a tuple $(P, T, s_0, E, L)$ where $P$ is
a finite set of {\em places}, $T$ is a finite set of {\em
  transitions}, $\mathit{init}$ is the set of initially
marked places, also referred to as the {\em initial
  marking}, $E$ is a finite set of {\em events}, and $L: T
\rightarrow E$ is a labeling function that maps each
transition $t \in T$ to an event $e \in E$.  For each
transition $t \in T$, its preset, denoted as $\bullet{t}
\subseteq P$, is the set of places connected to $t$, and its
postset, denoted as $t\bullet \subseteq P$, is the set of
places that $t$ is connected to.  A marking $s \subseteq P$
of a LPN is a subset of places marked with tokens, and it is
also referred to as a state of a LPN.  The initial marking
$\mathit{init}$ is also the initial state of the LPN.

%% with respect to information captured in the system
%% flows.  This requires recognizing signal events and mapping
%% the signal events or sequences of the signal events to flow
%% events appeared in system flow specifications.  For example,
%% a data write message may be a single event in a system flow,
%% however, such a flow event maybe implemented in hardware as
%% a sequence of events including a header, a number of data
%% word transfers and a tail.  This abstraction requires a
%% mapping relation from flow events to sequences of signal
%% events, and it is reasonable to assume that this relation is
%% available.

%% The trace interpretation takes a finite trace of flow events
%% resulting from the trace abstraction and a set of system
%% flows in LPNs $\vec{F}$, and generates a set of possible
%% system flow execution scenarios.  A \emph{flow execution
%%   scenario} is defined as $\{(F_{i,j}, s_{i,j})\}$ where in
%% each element $(F_{i,j}, s_{i,j})$, $F_{i,j}$ is the $j$th
%% activated instance of flow $F_i \in \vec{F}$, and $s_{i,j}$
%% is a state of $F_{i,j}$.  A flow execution scenario
%% indicates that at a certain point of SUD execution, what
%% types of flows and the number of instances of a particular
%% flow are activated and their corresponding current states.

%% This formalism has well defined
%% operational semantics, efficient analysis analysis
%% algorithms and and has been used widely in modeling and
%% analyzing communication protocols, concurrent programs, etc.

%--- LPN definition with levels
%Specifically, a labeled Petri-net (LPN) is a tuple $(P, T, M_0, V, E, L)$ where
%\begin{itemize}
%\item $P$ is a finite set of places,
%\item $T$ is a finite set of transitions,
%\item $M_0$ is the set of initially marked places, also referred to as the initial marking.
%\item $V$ is a finite set of auxiliary variables of integer type,
%\item $E$ is a finite set of events,
%\item $L$ is a labeling function that maps each transition $t \in T$ to a triple $(g, e, A)$ where
%	\begin{itemize}
%	\item $g$ is a predicate over $V$,
%	\item $e \in E$ is an event,
%	\item $A$ is a set of assignments to some variables in $V$.  
%	\end{itemize}
%\end{itemize}
%
%For each transition $t \in T$, its preset, denoted as $\bullet{t} \subseteq P$, is the set of places connected to $t$, and its postset, denoted as $t\bullet \subseteq P$, is the set of places that $t$ is connected to.  A marking of a LPN is a set of places marked with tokens, and a state of a LPN is $(M, \alpha)$ where $M$ is a marking and $\alpha$ is an assignment to all auxiliary variables in $V$.  
%
%The operational semantics of a LPN is defined by transition executions.  A transition can be executed after it is {\em enabled}.  A transition $t = (g, e, A)$ is enabled in a state $(M, \alpha)$ if every place in its preset is included in the marking, i.e. $\bullet{t} \subseteq M$, and the values of the auxiliary variables make the predicate $g$ be true, i.e. $\alpha \models g$.  Execution of $t$ results in a new state $(M', \alpha')$ such that 
%\[
%M' = (M - \bullet{t}) \cup t\bullet,
%\]
%and $\alpha'$ is a new assignment where the values of the auxiliary variables are updated with respect to the assignments $A$.  Furthermore, when $t$ is executed, the event $e$ is emitted.  SoC

%% Formally, an LPN is a tuple $(P, T, s_0, E, L)$ where $P$ is
%% a finite set of {\em places}, $T$ is a finite set of {\em
%%   transitions}, $\mathit{init}$ is the set of initially
%% marked places, also referred to as the {\em initial
%%   marking}, $E$ is a finite set of {\em events}, and $L: T
%% \rightarrow E$ is a labeling function that maps each
%% transition $t \in T$ to an event $e \in E$.  For each
%% transition $t \in T$, its preset, denoted as $\bullet{t}
%% \subseteq P$, is the set of places connected to $t$, and its
%% postset, denoted as $t\bullet \subseteq P$, is the set of
%% places that $t$ is connected to.  A marking $s \subseteq P$
%% of a LPN is a subset of places marked with tokens, and it is
%% also referred to as a state of a LPN.  The initial marking
%% $\mathit{init}$ is also the initial state of the LPN.

%% The operational semantics of a LPN is defined by transition executions.  A transition can be executed after it is {\em enabled}.  A transition $t \in T$ is enabled in a state $s$ if every place in its preset is included in the marking, i.e. $\bullet{t} \subseteq s$.  Execution of $t$ results in a new state $s'$ such that 
%% \[
%% s' = (s - \bullet{t}) \cup t\bullet,
%% \]
%% and the emission of event $e$ labeled for $t$.

%% The communication protocol in the BMPN shown
%% in~Fig.~\ref{flow-spec-ex}(a) is represented by the LPN
%% shown in~Figure~\ref{flow-spec-ex}(b).  In this and the
%% following figures for LPNs, the labeled circles denote
%% places, and the labeled boxes denote transitions.  Each
%% transition is labeled with its name and the associated
%% event.  Each event has a form of $\langle {\tt src, dest,
%%   cmd} \rangle$ where ${\tt cmd}$ is a command sent from a
%% source component ${\tt src}$ to a destination component
%% ${\tt dest}$. The solid black places without outgoing edges
%% are {\em terminals}, which indicate termination of protocols
%% represented by the LPNs.  The initial marking is
%% $\mathit{init} = \{p_1\}$.  In this LPN model, only the
%% communication portion of the BPMN specification is
%% represented while the computation portion is ignored.

% @COMMENT This example is for LPNs with guards.
%The above protocol specification can be made more precise by describing how component~2 would respond without using non-determinism.  For example, the system architect may wish to specify that component~2  responds with $\mathit{msg_2}$ to every sequence of two $\mathit{msg_1}$ for five such sequences in a row.  On the sixth such sequence, component~2 responds with $\mathit{msg_3}$.  The LPN modeling such protocol is shown in Figure~\ref{fig-flow-ex-2}.  An alternative without using auxiliary variables in the transition predicates is by unrolling the loops for the specified number of times; however this would result in a larger and more complex LPN.  
%
%\begin{figure}[htbp]
%\begin{center}
%\includegraphics[height=1.6in]{figures/flow-ex-2}
%\caption{Another example of a LPN representing the system flow in Fig.~\ref{fig-flow-ex-1} with a fixed number of message exchanges.}
%\label{fig-flow-ex-2}
%\end{center}
%\end{figure}

%=====

\section{Flow-Directed Trace Interpretation}

%% In a typical validation setting, the system under debug
%% (SUD) is executed in a test environment until it is
%% terminated by the test environment or the system crashes
%% due to a failure.  During the execution, a trace on a
%% small number of observable signals is streamed off the
%% chip for debugging.  Due to the limited observability and
%% inherent non-determinism in today's SoC designs, the
%% observed signal trace is difficult to understand, thus
%% providing limited values for debugging.  In this section,
%% we describe a trace analysis method where the observed
%% signal traces are interpreted at the level of system
%% flows.  In general, the trace analysis can offer
%% debuggers a structured view of communications among the
%% IP blocks during the SUD execution by deriving the types
%% and numbers of system flows activated during SUD
%% executions from the observed signal traces.

%% The trace analysis method consists of two steps:
%% \emph{trace abstraction} and \emph{interpretation}.  The
%% trace abstraction takes a signal trace observed during
%% the execution of the SUD, and abstracts it with respect
%% to information captured in the system flows.  This
%% requires recognizing signal events and mapping the signal
%% events or sequences of the signal events to flow events
%% appeared in system flow specifications.  For example, a
%% data write message may be a single event in a system
%% flow, however, such a flow event maybe implemented in
%% hardware as a sequence of events including a header, a
%% number of data word transfers and a tail.  This
%% abstraction requires a mapping relation from flow events
%% to sequences of signal events, and it is reasonable to
%% assume that this relation is available.

In this section we formalize the trace interpretation
problem in terms of labeled Petri-nets, and discuss our
algorithms to address the problem.  For pedagogical reasons,
here we assume full observability of all hardware signals
involved in the flow events.  In the next section we will
extend the approach to partial observability.

\medskip

\noindent 
{\bf Notations and formalization.}  We assume that the set
of system flows is provided as a collection ${\vec{F}}$ of
LPNs.  We define a {\em flow execution scenario} to be a
set $\{(F_{i,j}, s_{i,j})\}$ where $F_{i,j}$ is the
$j$th instance of flow $F_i \in {\vec{F}}$, and $s_{i,j}$ is
a state of $F_{i,j}$.  A flow execution scenario indicates
the set of protocols and the number of instances of a
particular protocol are activated and their corresponding
current states.  Since we assume full observability, we view
an {\em observed trace} $\rho = e_1e_2\ldots e_n$ is a
sequence of events.  Given an observed trace $\rho$, the
goal of of trace interpretation is to construct a set of
candidate flow execution scenarios whose execution can
create the sequence of events in $\rho$.  We call such
execution scenarios {\em compliant} with $\rho$.  Let
$\mathit{accept(F_{i,j}, s_{i,j}, e)}$ be a function that
determines if event $e$ can be emitted by $F_{i,j}$ in state
$s_{i,j}$.  Formally, $\mathit{accept(F_{i,j}, s_{i,j}, e)}$
returns $(F_{i,j}, s'_{i,j})$ where $s'_{i,j} = (s_{i,j} - \bullet t)
\cup t\bullet$ if there exists a transition
$t$ in $F_i$ such that $L(t) = e$ and $\bullet t \subseteq
s_{i,j}$.  It returns $\emptyset$ if no such $t$
exists in $F_i$.

\medskip

Given an obsered trace $\rho$ and the set $\vec{F}$ of LPNs,
Algorithm~\ref{algo:compliance} provides a basic procedure for
computing a set of compliant flow execution scenarios.
\begin{algorithm}[ht]
\DontPrintSemicolon
Create an empty scenario $scen$\;
$\mathit{Scen} = \{scen\}$\;
\ForEach {$h, \; 1 \leq h \leq n $} {
	$\mathit{found} \gets {\tt true}$ \;
	$Scen' = \emptyset$\;
	\ForEach {$scen \in Scen$} {
  		\ForEach {$(F_{i,j}, s_{i,j}) \in \mathit{scen_1}$} {
    		\If {$\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = (F_{i,j}, s'_{i,j})$} {
				Let $scen'$ be a copy of $scen$\;
      		$\mathit{scen'} \gets (\mathit{scen'} - (F_{i,j}, s_{i,j})) \cup (F_{i,j}, s'_{i,j})$\;
      		$\mathit{Scen'} \gets \mathit{scen'} \cup \mathit{Scen'}$\;
      		$\mathit{found} \gets {\tt false}$ \;
    		}
  		}
  		\ForEach {$F_i \in \vec{F}$} {
      	create a new instance $F_{i, j+1}$ \;
      	\If {$\mathit{accept}(F_{i,j+1},\mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$} {
    			Let $scen'$ be a copy of $scen$\;
				$\mathit{scen'} \gets \mathit{scen'} \cup (F_{i,j+1}, s'_{i,j+1})$ \;
				$\mathit{Scen'} \gets \mathit{scen'} \cup \mathit{Scen'}$\;
        		$\mathit{found} \gets {\tt false}$ \;
      	}
    	}
	}
  \If {$\mathit{found} == {\tt true}$} {
    \Return ${\tt Inconsistent}$\;
  }
  $Scen = Scen'$\;
}
\Return $\mathit{scen}$ \;
\caption{$\textsc{Check-Compliance}(\vec{F}, \, \rho)$}
\label{algo:compliance}
\end{algorithm}
The algorithm operates by keeping track (in variable
$\mathit{Scen}$) of a set of candidate flow execution scenarios
compliant with each prefix of $\rho$.  At each iteration,
for each event $e_h$ in the observed trace, we
update $Scen$ by either extending a member of
$\mathit{scen}$ or initiating a new protocol instance for 
each $scen \in Scen$ with respect to $e_h$ in every possible way.
If $e_h$ cannot be emitted by any existing or new flow instances, 
then we report that the trace is {\tt  inconsistent}, \ie, 
there is no possible interleaving of
the protocol instances from $\vec{F}$ that is compliant with
$\rho$.

To illustrate the basic idea of the trace analysis method, 
consider the system flow shown in Figure~Fig.~\ref{flow-spec-ex}(b).
Let $F_1$ denote such flow.  Suppose that a hardware system 
implements flow $F_1$, and the following flow trace is abstracted 
from an observed signal trace as a result of executing such system.
\[
	t_1\;t_2\;t_1\;t_2\;t_3\;t_3\;t_4\;t_5\;t_5\;t_4\ldots
\]  
where transition names in the LPN are used to represent the 
flow events in the trace.  The first four events results in
the following flow execution scenario
\[
	\{(F_{1,1}, \{p_3\}),~(F_{1,2}, \{p_3\})\}.
\]
For the first event $t_3$, it results in two execution scenarios 
below depending on which flow instance emits $t_3$.
\[
\begin{array}{l}
	\{(F_{1,1}, \{p_4\}),~(F_{1,2}, \{p_3\})\} \\
	\{(F_{1,1}, \{p_3\}),~(F_{1,2}, \{p_4\})\}.
\end{array}
\]
After handing the next event $t_3$, the above two execution scenarios
are reduced to the one as shown below.
\[
	\{(F_{1,1}, \{p_4\}),~(F_{1,2}, \{p_4\})\}.
\]
By following Algorithm~\ref{algo:compliance} to handle the remaining 
four events, the following execution scenario can be derived.
\[
	\{(F_{1,1}, \{p_5, p_6\}),~(F_{1,2}, \{p_5, p_6\})\}
\]
%% Given a trace of flow events $\rho = e_1e_2\ldots e_n$, the
%% trace interpretation algorithm starts with an empty set of
%% of flow execution scenario $\mathit{Scen} = \emptyset$.
%% Then, for each $e_h$ where $1 \leq h \leq n$ starting $h=1$,
%% and for each $\mathit{scen} \in \mathit{Scen}$, the
%% following two steps are performed.
%% \begin{description}
%% \item[Step 1]~~For each $(F_{i,j}, s_{i,j}) \in
%%   \mathit{scen}$, if $\mathit{accept}(F_{i,j}, s_{i,j}, e_h)
%%   = (F_{i,j}, s'_{i,j})$, create a new scenario
%%   $\mathit{scen}' = (\mathit{scen} - (F_{i,j}, s_{i,j}))
%%   \cup (F_{i,j}, s'_{i,j})$, which is added into
%%   $\mathit{Scen}'$.

%% \item[Step 2]~~For each $F_i \in \vec{F}$, create a new
%%   instance $F_{i, j+1}$.  If $\mathit{accept}(F_{i,j+1},
%%   \mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$,
%%   create a new scenario $\mathit{scen}' = \mathit{scen} \cup
%%   (F_{i,j+1}, s'_{i,j+1})$, which is added into
%%   $\mathit{Scen}'$.
%% \end{description}

%% After $e_h$ is processed, $Scen = Scen'$, and the above two
%% steps repeat for the next event $e_{h+1}$.

%% If every events in $\rho$ is successfully mapped to some
%% flow instance, this algorithm returns a set of flow
%% execution scenarios such that every flow instance is in its
%% terminal state.  On the other hand, inconsistent events can
%% also be encountered.  An event $e$ is \emph{inconsistent} if
%% for each flow execution scenario $\mathit{scen} \in
%% \mathit{Scen}$, the following two conditions hold.

%% \begin{enumerate}
%% \item For each $(F_{i,j}, s_{i,j}) \in \mathit{scen}$,
%%   $\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = \emptyset$,
%% \item For each $F_i \in \vec{F}$, $\mathit{accept}(F_{i},
%%   \mathit{init}_{i}, e_h) = \emptyset$.
%% \end{enumerate}

%% An inconsistent event is the one produced by SUD execution
%% but cannot be mapped to any flow instances no matter how the
%% trace prior to event $e$ is interpreted.  Inconsistent
%% events indicates possible causes of system failures.

%% Based on the above discussion, the trace interpretation
%% algorithm returns two pieces of information: 1) a set $G$ of
%% flow execution scenarios where every flow instance in every
%% scenario is in its terminal state, 2) a set $B$ of pairs,
%% each of which includes a set of flow execution scenarios and
%% an inconsistent event.  The set $B$ provides valuable
%% information for debuggers to root cause system failures.
%% One of these two sets can empty.  With the full
%% observability, the set $G$ includes a single flow execution
%% scenario derived for a trace.  In reality, it is always the
%% case that the SUD is only partially observable.  Therefore,
%% due to the lack of information for precise interpretation, a
%% set of flow execution scenarios is typically derived for a
%% given trace as the result of the trace analysis.

%% \red{\bf The following illustration may not be needed (from here to the end of this subsection) if seciton IV is kept.  Or we keep the illustration below but remove section IV which uses 3 pages.}

%% To illustrate the basic idea of the trace analysis method, consider the system flow shown in Figure~\ref{fig-flow-ex-1}.  Let $F_1$ denote such flow.  Suppose that a hardware system implements flow $F_1$, and the following trace of flow events is abstracted from an observed signal trace as a result of executing such system.     
%% \[
%% \mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_3\;\ldots
%% \]  
%% This trace is interpreted from the first event to the last in order to derive all possible flow execution scenarios.  At the beginning, the first event $\mathit{msg}_1$ is processed.  According to the flow specification $F_1$, we know that one instance of such flow $F_1$, $F_{1,1}$,  is activated by the SUD as $\mathit{accept}(F_{1,1}, init_1, \mathit{msg}_1) = (F_{1,1}, \{p_2\})$ where $\mathit{init} = \{p_1\}$ is the initial state of $F_1$.  As the result, the flow execution scenario after interpreting the first event $\mathit{msg_1}$ is $\{(F_{1,1}, \{p_2\})\}$.

%% Next, the second $\mathit{msg}_1$ is interpreted on both scenarios   This event could be the result of two possible cases.  In the first case, this event is the result of the continuing execution of $F_{1,1}$ as $\mathit{accept}(F_{1,1}, \{p_2\}, \mathit{msg}_1) = (F_{1,1}, \{p_3\})$.  In the second case, the system may activate another instance of $F_1$, $F_{1,2}$ such that the second event $\mathit{msg_1}$ is a result of executing this new instance.  Therefore, the interpretation of the first two events $\mathit{msg}_1$ leads to two flow execution scenarios as shown below.
%% \[
%% \label{tr-analy-s1}\tag{1}
%% \begin{array}{cl}
%% 1. & \{(F_{1,1}, \{p_3\})\},\\
%% 2. & \{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_2\})\} 
%% \end{array}
%% \]

%% Now, consider the third $\mathit{msg}_1$ for each of the two scenario derived in~(\ref{tr-analy-s1}).    For the execution scenario~1, $F_{1,1}$ is not able to accept $\mathit{msg}_1$ as it is in state $\{p_3\}$.  On the other hand, this event could be the result of activation of a new flow instance.  Therefore, this execution scenario can be revised accordingly as 
%% \[
%% \label{tr-analy-s2}\tag{2}
%% \{(F_{1,1}, \{p_3\}),(F_{1,2},\{p_2\})\}.
%% \]   
%% For the execution scenario~2, event $\mathit{msg}_1$ could be the result of continuing execution of $F_{1,1}$ or $F_{1,2}$, or it could be a result of activation of a new flow instance.  Therefore,  three new execution scenarios can be derived as shown below for this event.
%% \[
%% \label{tr-analy-s3}\tag{3}
%% \begin{array}{l}
%% \{(F_{1,1}, \{p_3\}), (F_{1,2}, \{p_2\})\},\\
%% \{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_3\})\}, \\
%% \{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_2\}), (F_{1,3}, \{p_2\})\}
%% \end{array}
%% \]
%% Since the flow execution scenario in~(\ref{tr-analy-s2}) already exists in~(\ref{tr-analy-s3}), the three flow execution scenarios shown in~(\ref{tr-analy-s3}) is the result from the interpretation of the first three events $\mathit{msg}_1$.

%% The next flow event in the trace $\mathit{msg_2}$ is analyzed for the flow execution scenarios as shown in~(\ref{tr-analy-s3}).   For the first scenario $\{(F_{1,1}, \{p_3\}), (F_{1,2}, \{p_2\})\}$, event $\mathit{msg_2}$ can only be the result from executing $F_{1,1}$ as $\mathit{accept}(F_{1,1}, \{p_3\}, \mathit{msg_2}) = (F_{1,1}, \{p_1\})$.   Based on the same reasoning, this event can only be the result from executing $F_{1,2}$ in the second scenario, and it moves $F_{1,2}$ to a new state $\{p_3\}$ too.  The interesting case is the third scenario where none of the flow instances can allow $\mathit{msg_2}$ to happen.  This is due to the fact that the flow must be in $\{p_3\}$ for $\mathit{msg_2}$ to happen.  This indicates the third system execution scenario is impossible for the prefix of the flow event trace upto $\mathit{msg_2}$, therefore this scenario is ignored from further analysis.  After analyzing event $\mathit{msg_2}$, the updated system execution scenarios are shown below.
%% \[
%% \begin{array}{l}
%% \{(F_{1,1}, \{p_1\}), (F_{1,2}, \{p_2\})\},\ \{(F_{1,1}, \{p2\}), (F_{1,2}, \{p_1\})\}.
%% \end{array}
%% \]

%% The last flow event in the trace is $\mathit{msg}_3$.  Considering the above two possible system executions, neither can allow this event to be produced as none of the flow instances in both system executions is in state $\{p_3\}$.  What this means is that the system does {\em not} implement the flow specification correctly as it produces something not allowed by the specification.  By examining the system executions right before the ``buggy'' event, debuggers may gain more information on when and where the problem might be.  The trace interpretation algorithm adds these two scenarios along with the fifth event $\mathit{msg_3}$ into $B$, and returns it for debuggers to analyze.

%---
\section{Trace Analysis with Partial Observability}

The trace interpretation discussion in the preceding section
assumed full observability.  In this section, we discuss how
to adapt the method to deal with silicon traces under
limited (partial) obsrvability.

%% In hardware that implements a given system flow
%% specification, a flow event is defined as an event or a
%% sequence of events on a set of signals.  Due to the limited
%% number of pins on the boundary of chips available for
%% observation, only a small fraction of system signals can be
%% observed during debug.  In this section, we discuss how the
%% trace analysis method presented above can be adapted to deal
%% with signal traces of partial observability.

In general, a signal trace of partial observability corresponding a set of traces of flow events due to the ambiguous interpretation of signal events.  In the following, we discuss two cases for trace abstraction on partial observability: mapping a single signal event to a flow event or mapping a sequence of signal events to a flow event.  A signal event is defined as a state on or an assignment to a set of signals.

Hereafter, the term {\em flow traces} is used to refer to traces of flow events. 
Consider the following example for the first case.  Suppose that there are three flow events: $e_1$, $e_2$, and $e_3$, which are implemented in hardware by the signal events shown in the list below.  We use Boolean expressions to represent signal events for the discussion.
\[
\begin{array}{cl}
e_1: & abc\\
e_2: & \bar{a}bc\\
e_3: & a\bar{b}c
\end{array}
\] 
Now suppose that only signals $b$ and $c$ are observable, and a signal trace of this partial observability is obtained below.
\[
bc\ bc \ \bar{b}c
\]
During the trace abstraction step, the first two signal events $bc$ can be mapped to $\{e_1, e_2\}$ since $a$ is not observable, and the last one $b'c$ is mapped to $\{e_3\}$.  Therefore, this signal trace is abstracted to four flow traces, $\{e_1, e_2\} \times \{e_1, e_2\} \times \{e_3\}$.

Next, we consider the case where a flow event is mapped from a sequence of signal events.  Now suppose that two other flow events are implemented by sequences of signal events as defined in the list below.
\[
\begin{array}{cl}
e_4: & abc\ \bar{a}bc\\
e_5: & abc\ abc\ abc\ \bar{a}bc\\
\end{array}
\] 
Given an observed trace of the same observability shown below
\[
bc\ bc\ bc \ bc,
\]
it is abstracted to the following flow traces.
\[
e_4e_4,\; \textvisiblespace e_4 \textvisiblespace,\;  e_4 \textvisiblespace \textvisiblespace,\;  \textvisiblespace \textvisiblespace e_4, \; e_5
\] 
where $\textvisiblespace$ denotes signal events that are not mapped to any flow events.  Note that the above abstraction leads to three distinct flow traces as the middle three correspond to the same trace of flow events.

From the above discussions, it can be seen that a signal trace of partial observability is generally mapped to a set of flow traces.  As the the number of signals available for observation becomes smaller, the number of flow event traces corresponding to a signal trace as a result of the abstraction can be enormous.  This phenomenon can increase the complexity of the trace interpretation as it can cause the number of possible flow execution scenarios generated during the analysis to explode.   Possible solutions to address this issue include better trace signal selection for observation and assistance from debuggers' insights.  Trace signal selection itself is an important and difficult subject, and a detailed discussion of it is out of scope of this paper.   Next, we briefly describe how the debuggers' insights of a system's architecture can help to address the complexity issue in the trace analysis.

\subsection{Inputs from Validators}

During the trace interpretation, the number of intermediate flow execution scenarios may become too large due to ambiguous interpretation from signal events to flow events or from flow events to flow execution scenarios.  The explosion of the intermediate results can significantly slow down the performance of the trace analysis.  To address this problem, the debuggers can use their insights and understanding of the SUD to trim the total number of possible system executions derived from the trace analysis.  When a SUD is validated, a debugger is assumed to have deep knowledge about the system's architecture and microarchitecture, and how the test environment affects the system executions.  For instance, he/she may have knowledge that in a test environment, the maximal number of instances of a flow can be activated by the SUD, or a flow can be activated after certain other flows have terminated, etc.  These debugging insights can be encoded as constraints into the trace analysis method, which can then be used to eliminate a large number of flow execution scenarios that violate these constraints during the trace interpretation step.

This approach can be flexible in that it allows a debugger to analyze the observed traces in a trail-and-error manner if the precise knowledge of the system (micro-)architecture is hard to come by.  For instance, the debugger might initially make a very restricted assumption on how the SUD executes a flow specification, and these assumptions can potentially lead to an empty set of flow execution scenarios.  Depending on which of these assumptions triggered during the trace interpretation step, the debugger can study these assumptions more carefully, and relax some or all of them for the next run of analysis.  This iteration can be repeated as many times as necessary until some results deemed meaningful are produced.

Alternatively, if all derived execution scenarios seem to be plausible, the implication that a debugger may draw from this result is that the failure may be independent of the flows being observed.  Therefore, the testing environment can be adjusted in order for a different part or different behavior of the SUD to be observed.  This idea, closely related to trace signal selection, is critical for post-silicon validation, and a detailed discussion can only be presented in a separate paper.

%%-- Case study I
%\input{case1}

%
%--
%
\section{Case Study}

\begin{figure} 
\centerline{
\includegraphics[width=3.3in]{figures/Fig4.png}}
\caption{SoC platform structure.}
\label{SoC}
\end{figure}

In order to find out the efficiency of the trace analysis method for more realistic examples, in this case study, a more detailed transaction level model of a SoC is constructed within the GEM5 environment~\cite{Binkert2011}.  This SoC model consists of two ARM Cortex-A9 cores, each of which contains two separate 16KB data and instruction caches.  The caches are connected to a 1GB memory through a memory bus model.  The architecture of this SoC model is shown in Fig.~\ref{SoC}.

In this model, components communicate with each other by sending and receiving various request and response messages.  In order to observe and trace communications occurring inside this model during execution, monitors are attached to links connecting the components. These monitors record the messages flowing through the links they are attached to, and store them into output trace files.

%By combining the informations from all of the nine communication monitors, required communication traces are obtained. As a virtualized SoC platform, GEM5 has three types of request: timing, atomic and functional. Timing request include the modeling of queuing delay and resource contention. Atomic request is a faster than timing request with no delay.  Functional request is used for loading binaries, examining/changing variables in the simulated system, and so on. In our case, we specify all our request to be timed as it's the most detailed one.However, some of the system initiation is still atomic and it's not related to our research, so we only took the messages that are timing request from our collected data.

%\subsection{Flow Specification}

For this model, we consider the flow specifications describing the cache coherence protocols supported in GEM5 that is used to build the model in Fig~\ref{SoC}.  These flow specifications describe data/instruction read operations and data write operations initiated from CPUs.  Three such flows describe the cache coherent protocols for each CPU.  As there are two CPUs in this model, there are actually six such flows considered for this model.  These flow specifications are derived from the cache coherence protocols supported in GEM5\footnote{The GEM5 cache coherence protocols can be found at \url{http://www.gem5.org/docs/html/gem5MemorySystem.html}.}.  More detailed information about the flow specifications used in this case study will be provided as a supplement document if this paper is accepted for publication.   
%More detailed information about these protocols and the flow specifications can be found online\footnote{\url{https://github.com/cao2/paper/blob/master/Flow\%20Specification.pdf}} to meet the page limit.

%In the following section, a brief overview for the implemented instances of these flows are given. These flow specifications capture how messages are exchanged for different use cases. In this model, a message is defined with the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform. 
%\begin{figure} 
%\centerline{
%\includegraphics[width=3.7In]{figures/Fig5.png}}
%
%{\footnotesize
%\[
%\begin{array}{llll}
%msg_0: (&\mbox{ CPU1},&\mbox{writeReq},&\mbox{cache1   })\\       
%msg_1: (&\mbox{ cache1},&\mbox{ readExreq },&\mbox{Bus     })\\        
%msg_2: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2 })\\  
%msg_3: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2         })\\   
%msg_4: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2           })\\  
%msg_5: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2 })\\  
%msg_6: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce1       })\\     
%msg_7: (&\mbox{ cache1},&\mbox{readExreq},&\mbox{cpu1           })\\  
%msg_8: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ Memory })\\  
%msg_9: (&\mbox{ true                                          })\\  
%msg_{10}: (&\mbox{ Memory},&\mbox{ readExres},&\mbox{ Bus        })\\  
%msg_{11}: (&\mbox{ cache2},&\mbox{ readExres},&\mbox{ Bus })\\  
%msg_{12}: (&\mbox{ Bus},&\mbox{ readExres},&\mbox{ cache1        })\\  
%msg_{13}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1         })\\  
%msg_{14}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1 })\\  
%msg_{15}: (&\mbox{ cache1},&\mbox{UpgradeReq             })\\  
%msg_{16}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ cahce2      })\\   
%msg_{17}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ Memory })\\  
%msg_{18}: (&\mbox{ cache2},&\mbox{UpgradeRes},&\mbox{ Bus     })\\  
%msg_{19}: (&\mbox{ Bus},&\mbox{ UpgradeRes},&\mbox{ cache1      })\\  
%msg_{20}: (&\mbox{ cache1},&\mbox{ WriteRes},&\mbox{ CPU1 })\\  
%\end{array}
%\]}
%\caption{Flow specification ($F_1$) of a cache coherent write operation initiated from CPU1}
%\label{write-flow}
%\end{figure}
%
%The LPN as shown in Fig.~\ref{write-flow} specifies a system flow where CPU1 initiates a memory write operation. In this flow, CPU1 initiates a memory write request to L1 Cache. Next, depends on whether the required data is included in cache1, the cache1 will generate three possible responses. First, the cache1 will send read exclusive request message msg2 to interconnect if data is not present in cache.; Or if the data is shared by CPU0, it will send upgrade message msg16 to interconnect to disable CPU0's ownership of this block of data, else if CPU1 has exclusive right of required data, cache1 will perform the write operation and sent an response message msg15 to CPU1. Afterwards, interconnect will sent request to all connected component (data cache and instruction cache of each CPU and memory). Once the interconnect obtains the response from either CPU0 or memory, it will generate a response message to cache1, then cache1 will generate a write response message msg14 (or msg 21 ) to CPU1. The flow is symmetric for CPU2.
%
%
%\begin{figure} 
%\centerline{
%\includegraphics[width=4in]{figures/Fih6.png}}
%
%{\footnotesize
%\[
%\begin{array}{llll}
%t0: (&\mbox{ CPU1},&\mbox{ReadReq},&\mbox{cache1  })\\                   
%t1: (&\mbox{ cache1},&\mbox{ StoreCondreq },&\mbox{Bus })\\           
%t2: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2 })\\
%t3: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2       })\\      
%t4: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2           })\\ 
%t5: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2 })\\
%t6: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce1     })\\       
%t7: (&\mbox{ cache1},&\mbox{StoreCondreq},&\mbox{cpu1           })\\ 
%t8: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ Memory })\\
%t9: (&\mbox{ true                                        })\\
%t10: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus            })\\    
%t11: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus })\\
%t12: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1      })\\            
%t13: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1          })\\  
%t14: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
%t21: (&\mbox{ cache1},&\mbox{ LoadLockedreq },&\mbox{Bus })\\     
%t22: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\
%t23: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2 })\\
%t24: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\ 
%t25: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2     })\\
%t26: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce1 })\\
%t27: (&\mbox{ cache1},&\mbox{LoadLockedreq},&\mbox{cpu1       })\\
%t28: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ Memory     })\\
%t29: (&\mbox{ true })\\
%t30: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus       })\\     
%t31: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus    })\\
%t32: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1 })\\
%t33: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
%\end{array}
%\]}
%\caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1}
%\label{read-flow}
%
%\end{figure}
%The LPN specification as shown in Fig.~\ref{read-flow} captures the system flow where CPU1 initiate a memory read operation. Similar to memory write, CPU1 will first initiate a memory read request message msg1 to cache1. Cache1 can generate three possible responses. First, if the requested data is not in cache1, cache1 will generate an storeCondReq message msg2 to interconnect asking for data. Second, if the requested data is inside of cache1 but it's shared, cache1 will generate a upgrade request message msg22 to make sure it has the newest data; Last,  if cache1 has and data and it's not shared, it will generate a read response message msg15 to CPU1. Afterwards, interconnect will send corresponding request to all of its connected components. When interconnect receives the response, it will generate response message to cache1. Then cache1 will generate the read response message msg14 or msg34 to CPU1. This specification is also symmetric for CPU2.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Results and Discussions}

Two simple programs are written, one for each CPU.  These simple programs read numbers from a file, perform some operations on these numbers, and store the results back to the file.  How GEM5 supports shared memory multi-threaded program execution is unclear.  Therefore, there is no data shared in both caches in this test.  Furthermore, GEM5 does not support true concurrency.  When there are two programs running on the CPUs, GEM5 alternates the executions between the two CPUs.  To simulate asynchronous concurrency with the interleaving semantics, those two simple programs are instrumented with pseudo-blocking commands, one placed before each statement.  A pseudo blocking command includes a random number generator that returns either $0$ or $1$ and a loop that only exits when the returned random number is $0$.

%We produce a result file including all of the communication messages between every components from two simple programs running simultaneously on each CPU.  The program assigned to CPU1 read one file three times for one letter and writing one letter for three times to the same file. CPU2's program will do the same read and write functionalities to the same file, only difference is that CPU2 will write first and then read. One thing we should know about GEM5 is that even when we run these two programs concurrently, it will attempt to produce a concurrent result. The data we collected shows that it's not the real nondeterministic concurrency. What GEM5 did was it allowed two CPUs to execute its own instruction in turn, therefore the order was deterministic. Therefore, no matter how many times we ran the program, it produced the same result. We tried other virtual SoC platform softwares, and this was the best nondeterministic concurrency we can get. Even if this is slightly off with what the real chip should work, it still surve our purpose of testing the correctness of our algorithm.

After the SoC model finishes executing the program, there are totally $343581$ messages collected in the trace file.  Not all of the messages are relevant to the flow specification as many are used by GEM5 to initialize its simulation environment.  After removing those irrelevant messages, the number of messages in the trace file is to reduced to $121138$.

The time taken to remove the irrelevant messages from the trace is negligible.  The total runtime and the peak memory usage for the trace analysis algorithm to finish the reduced trace are 3 seconds and 12MB, respectively.  From the trace, Table~\ref{table-case-2} shows the number of instances extracted for the six flows describing cache coherent read/write operations initiated from both CPUs. 
\begin{table}[tb]
\caption{The number of flow instances derived by the trace analysis with the full observability.}
\begin{center}
\begin{tabular}{|l|c|}
\hline
Flows & $\#$Instances \\
\hline
\hline
CPU1 Data Read			&  $17582$\\
CPU1 Instruction Read		&  $4002$\\
CPU1 Write				&  $3370$\\
\hline
CPU2 Data Read			&  $17386$\\
CPU2 Instruction Read		&  $3955$\\
CPU2 Write				&  $3308$\\
\hline
\end{tabular}
\end{center}
\label{table-case-2}
\end{table}%

To take the partial observability into account, the four monitors attached to the links between two CPUs and their caches are disabled. Then, the trace is generated by the remaining five monitors from the SoC model executing the same program. After removing all non-observable messages, the trace only contains 15089 messages. The numbers of the flow instances extracted by the trace analysis method are shown in Table~\ref{table-par-obs}.  From these results, the numbers of the flow instances are dropped significantly compared to the results extracted from the trace with the full observability as shown in Table~\ref{table-case-2}. This difference is due to that some communications occurred in the system when executing the program involve the CPUs and their corresponding caches only, and the traffic on the links between the CPUs and their corresponding caches is not observable. Therefore, the instances of the flow specifications characterizing these communications do not exist in the trace. In other words, all extracted flow instances in Table~\ref{table-par-obs} characterize the communications that pass through the memory bus in the system model.  The runtime and memory usage is similar to that for analyzing the trace of the full observability as shown above.

\begin{table}[tb]
\caption{The number of flow instances derived by the trace analysis with certain monitors disabled.}
\begin{center}
\begin{tabular}{|l|c|}
\hline
Flows & $\#$Instances \\
\hline
\hline
CPU1 Data Read			&  $829$\\
CPU1 Instruction Read		&  $169$\\
CPU1 Write				&  $82$\\
\hline
CPU2 Data Read			&  $803$\\
CPU2 Instruction Read		&  $190$\\
CPU2 Write				&  $83$\\
\hline
\end{tabular}
\end{center}
\label{table-par-obs}
\end{table}%

In the third experiment, further partial observability is taken into consideration.  In this experiment, only the five links involving the memory bus are still considered.  However, an assumption is made that all events passing the same link are not distinguishable due to the limitation of the observability.  More specifically, the monitors are modified such that whenever an event is captured on one of the links, it dumps a set of events passing through the same link into the trace file.  Therefore, each line of the trace file corresponds to a set of events.  After applying the trace analysis to this trace,  a total of 13944 flow execution scenarios are extracted.    This large number, compared to the number of extracted execution scenarios shown in the Table~\ref{table-case-2} and \ref{table-par-obs}, is due to the ambiguous interpretation of the events with limited observability.

The whole experiment takes about $15$ minutes and $420$~MB to finish, significantly higher than the numbers for analyzing traces where there is no ambiguity in the observed events.  This is due to the fact that a trace of ambiguous events is in fact a set of traces of original events, which lead to large numbers of execution scenarios either during or at the end of the analysis.  In this experiment, the peak number of execution scenarios during the analysis process is $70384$.   Compared to the final number, many of the intermediate execution scenarios are invalid, and removed eventually.  However, controlling the number of intermediate execution scenarios during the trace analysis is critical in order for the analysis to be tractable.  Here, insights from validators can help.

As shown above, the ambiguous interpretation of events can lead to large numbers of intermediate and final execution scenarios, which not only make the trace analysis more time consuming but also make it difficult to gain insightful understanding from the derived execution scenarios.  Careful selection of what to observe may have big impact on results from the trace analysis.  In this last experiment, we relax the assumption made in the previous experiment such that the events passing each link are partitioned into two groups, one for read operations and one for write operations.  Similar to the assumption made in the previous experiment, events in the same group are assumed to be non-distinguishable.  The monitors are modified accordingly such that they output all events in the same group into the trace file if an event from that group is captured.  After the trace analysis on this new partially observed trace is finished, only one execution scenario is derived where the distribution of the numbers of flow instances is the same as those shown in Table~\ref{table-par-obs}.  The peak number of execution scenarios encountered during the trace analysis is $4$.  The total runtime is about $1$~second.  Compared to the results from the previous experiment, the precision and the performance of the trace analysis are improved dramatically as a result of careful selection of observable signals.

%=====
\section{Related Work}

Our work is closely related to communication-centric and transaction based debug.  An early pioneering work is described in~\cite{Goossens2007NOCS}, which advocates the focus on observing activities on the interconnect network among IP blocks, and mapping these activities to transactions for better correlation between computations and communications.  Therefore, the communication transactions, as a result of software execution, provide an interface between computation and communication, and facilitate  system-level debug.  This work is extended in~\cite{Vermeulen2009VLSI-DAT,Goossens2009DATE}.  However, this line of work is focused on the network-on-chip (NoC) architecture for interconnect using the run/stop debug control method.

A similar transaction-based debug approach is presented in~\cite{Gharehbaghi2012ISQED}.  Furthermore, it proposes an automated extraction of state machines at transaction level from  high level design models.  From an observed failure trace, it performs backtracking on this transaction level state machine to derive a set of transaction traces that lead to the observed failure state.  In the subsequent step, bounded model checking with the constraints on the internal variables is used to refine the set of transaction traces to remove the infeasible traces.  This approach requires user inputs to identify impossible transaction sequences, and may not find the states causing the failure if the transaction traces leading to the observed failure state is long.  Backtracking from the observed failure state requires pre-image computation, which can be computationally expensive.  A transaction-based online debug approach is proposed in~\cite{Dehbash2014} to address these issues.  This approach utilizes a transaction debug pattern specification language~\cite{Gharehbaghi2009ICCD} to define properties that transactions should meet.  These transaction properties are checked at runtime by programming debug units in the on-chip debug infrastructure, and the system can be stopped shortly after a violation is detected for any one of those properties.  In this sense, it can be viewed as the hardware assertion approaches in~\cite{Boule2007ISQED} elevated to the transaction level.

In~\cite{Singerman2011DAC}, a coherent workflow is described where the result from the pre-silicon validation stage can be carried over to the post-silicon stage to improve efficiency and productivity of post-silicon debug.  This workflow is centered on a repository of system events and simple transactions defined by architects and IP designers.  It spans across a wide spectrum of the post-silicon validation including DFx instrumentation, test generation, coverage, and debug.  The DFx instruments are automatically inserted into the design RTL code driven by the defined transactions.  This instrumentation is optimized for making a large set of events and transactions observable.   Test generation is also optimized to generate only the necessary but sufficient tests to allow all defined transactions to be exercised.   Moreover, coverage for post-silicon validation is now defined at the abstract level of events and transactions rather than the raw signals, and thus can be evaluated more efficiently.  In~\cite{Abarbanel2014DAC}, a model at an even higher-level of abstraction, {\em flows}, is proposed.  Flows are used to specify more sophisticated cross-IP transactions such as power management, security, etc, and to facilitate reuse of the efforts of the architectural analysis to check HW/SW implementations.

%
%--
\section{Conclusion}

This paper presents a trace analysis based method for post-silicon validation by interpreting observed raw signal traces at the level of system flow specifications.  The derived flow execution scenarios provide more structured information on system operations, which is more understandable to system validators.   This information can help to locate design defects more easily, and also provides a measurement of validation coverage.

Due to partial observability, this approach may derive a large number of different flow execution scenarios for a given signal trace.  Insights from system validators can help to eliminate some false scenarios due to the partial observability.  An interesting future direction is formalization of the validators' insights using temporal logic on flows so that the validators can express their intents more precisely and concisely.

The trace analysis approach presented in this paper needs to be iterated with different observations selected in different iterations in order to eliminate the false scenarios and to root cause system failures as quickly as possible.  The observation selection and stitching signal traces of different observations together for the above goal will also be pursued in the future.

% conference papers do not normally have an appendix

% use section* for acknowledgement
%\section*{Acknowledgment}
%
%
%The authors would like to thank...
%

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{SoC}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
%
%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
%\end{thebibliography}

% that's all folks
\end{document}